```{r setup, include=FALSE}

knitr::opts_chunk$set(cache=FALSE)
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(include=TRUE)
knitr::opts_chunk$set(results="asis")
knitr::opts_chunk$set(fig.width=12)
knitr::opts_chunk$set(fig.height=14)
prjpath <- "~/work/learn/competitions/kaggle/rossmann/"
datapath <- paste(prjpath, "data/", sep="")
analpath <- paste(prjpath, "analysis/", sep="")
rcodepath <- paste(analpath, "Rcode/", sep="")
setwd(analpath)

```


```{r loadLibsAndSource, include=FALSE}

reqdpkgs <- c("lattice", "latticeExtra", "ggplot2", "reshape2",
              "plyr",  "lubridate", "xgboost",  "corrplot",
              "caret", "randomForest")
lapply(reqdpkgs, library, character.only=TRUE)

```

#Introduction


We will briefly analyse the Rossmann dataset to identify any
pecularities we need to notice while loading the data.


From Kaggle we have the following information on the data-set,

Most of the fields are self-explanatory. The following are descriptions for those that aren't.

1. Id - an Id that represents a (Store, Date) duple within the test set
2. Store - a unique Id for each store
3. Sales - the turnover for any given day (this is what you are predicting)
4. Customers - the number of customers on a given day
5. Open - an indicator for whether the store was open: 0 = closed, 1 = open
6. StateHoliday - indicates a state holiday. Normally all stores, with few exceptions, are closed on state holidays. Note that all schools are closed on public holidays and weekends. a = public holiday, b = Easter holiday, c = Christmas, 0 = None
7. SchoolHoliday - indicates if the (Store, Date) was affected by the closure of public schools
8. StoreType - differentiates between 4 different store models: a, b, c, d
9. Assortment - describes an assortment level: a = basic, b = extra, c = extended
10. CompetitionDistance - distance in meters to the nearest competitor store
11. CompetitionOpenSince[Month/Year] - gives the approximate year and month of the time the nearest competitor was opened
12. Promo - indicates whether a store is running a promo on that day
13. Promo2 - Promo2 is a continuing and consecutive promotion for some stores: 0 = store is not participating, 1 = store is participating
14. Promo2Since[Year/Week] - describes the year and calendar week when the store started participating in Promo2
15. PromoInterval - describes the consecutive intervals Promo2 is started, naming the months the promotion is started anew. E.g. "Feb,May,Aug,Nov" means each round starts in February, May, August, November of any given year for that store

The primary data is in the file train.csv, and secondary data in
store.csv,

```{r loaddata}

training <- read.csv("../data/train.csv")
testing <- read.csv("../data/test.csv")
store <- read.csv("../data/store.csv")

```

```{r processdates}

training$Date <- ymd(training$Date)
training$Year <- year(training$Date)
training$Month <- month(training$Date)
training$Week <- week(training$Date)
training$Mday <- mday(training$Date)
training$Wday <- wday(training$Date)

preds.dates <- c("Year", "Month", "Week", "Mday", "Wday")

```

#Store data

Other than identifying the store, the dataset training contains
columns that depend on the day that the row identifies. Further
details on the stores are specified in the data-set store. For any
prediction we will need to include the information from store columns
into the training data. We cannot directly copy the infomration from
*store*. We will have to interpret the *store* columns before we can
use them.

##Store Type and Assortmnet

The store type data type and can be included directly,

```{r storetype}

store <- store[order(store$Store),]
row.names(store) <- store$Store
training$StoreType <- store$StoreType[training$Store]
training$Assortment <- store$Assortment[training$Store]

```


##Competition

A distance in space, and a time of start of the competition is
provided. Using the time of the competition's start we can determine
if there was competition on the day of the sale.

```{r existsCompetition}

store$CompDate <- with(store,
                       ymd(paste(CompetitionOpenSinceYear,
                                 CompetitionOpenSinceMonth,
                                 1, sep="-")))
training$CompDate <- store$CompDate[training$Store]
training$IsCompetition <- training$Date > training$CompDate

```

We can also compute the duration the competition has existed, to model
situations in which the competition pulls away regular customers. In
reverse the store itself could have come up after the competition was
already there, and could pull customers away from the competition.

To determine what store to go to a customer should know the existance
of the stores. If a competition has existed forever, a newly opened
store should attract more customers over time, as should a newly
opened competition pull away the store's customers. What is the first
data of sales for each of the stores?

```{r firstsale}

firstsale <- ddply(training, "Store", summarise, Open = min(Date))

print('The stores have data beginning since one of the following dates')
print(table(firstsale$Open))

```
All the Rossmann stores have almost the same first sale date. For our
models we will assume that these stores started business on
2013-01-01.

We can count the number of days from the start day of the competition
to the day of the sale.

```{r compdur}

training$CompDuration <- with(training,
                              Date - CompDate)/edays(1)

```

##Promotions

There are two kinds of promotions. The variable *Promo* tells us if a
the store has a promotion on a particular day, and the variables
associated with *Promo2* tell us about a store's participation in a
continuing and consecutive promotion, which we can directly include in
*training*. The second promotion being
continuous, it cannot be readily expressed as a function of time. All
we know about the second promotion (Promo2) has to be derived from the
variables *Promo2SinceWeek*, *Promo2SinceYear*, and
*PromoInterval*. From the descriptions of these variables we can
assume that once started from the week and year provided in the data,
Promo2 is continuous. To reinforce, or to remind the customers of this
promotion, the store restarts the promotion during the months provided
in *PromoInterval*. We can use the data in the raw-form as it is, and
we can also derive a variable for each day which tells us the number of
days since Promo2 has been restarted.

We can encode variable *PromoInterval* with an integer. The group
Jan,Apr,Jul,Oct will then be 1, Feb,May,Aug,Nov will be 2, and
Mar,Jun,Sept,Dec will be 3. We do not have to write much code for
this, and can use the levels assigned by R. Notice that the empty
string, for stores not participating in Promo2, has been assigned
level 1. We will relevel *PromoInterval*, and then convert it to an integer,

```{r promo2dayssincej}

training$Promo2 <- store$Promo2[training$Store]

store$PromoInterval <- as.integer(factor(store$PromoInterval,
                                         c("", "Jan,Apr,Jul,Oct",
                                           "Feb,May,Aug,Nov",
                                           "Mar,Jun,Sep,Dec")
                                         )
                                  ) - 1

grp.mon <- rep(c(1,2,3), 4)
mon.grp <- as.data.frame(do.call(rbind,
                                 lapply(0:3, function(g) {
                                     if (g == 0) rep(NA, 4)
                                     else which(grp.mon == g)
                                 }))
                         )

Promo2RestartMonth <- mon.grp[store$PromoInterval + 1, ]
names(Promo2RestartMonth) <- paste("Promo2RestartMonth", 1:4, sep=".")

Promo2SinceRestart <- foreach(i=1:4, .combine = 'cbind') %do% {
    prd <- training$Date
    month(prd) <- Promo2RestartMonth[training$Store, i]
    day(prd) <- 1
    prd <- (training$Date - prd)/edays(1)
    prd[prd < 0] <- NA
    prd
}

training$Promo2DurationSinceRestart <- apply(Promo2SinceRestart, 1,
                                                   function(xs) {
                                                       min(xs,
                                                           na.rm=TRUE)
                                                   })

```

We have variables that tell us if a store is participating in Promo2
or not. We can check what the value of durations since promo2 restart
for these stores is (as a check on our procedure).

```{r testpromo2dur}

table(training$Promo2DurationSinceRestart[ training$Promo2 == 0])

```

As we can see, the duration since the restart of *Promo2* is infinite
for all stores that do not participate in *Promo2*, as should be.

# Exploratory Data Analysis
Since we know the column names we can engage in some insightful
explorations into the data. First we plot to see how the Sales scale
with the number of Customers.

```{r plotSalesCustomers}

sc <- training[, c("Sales", "Customers")]
sc <- sc[order(sc$Customers),]
xyplot(log10(Sales) ~ log10(Customers),
       data=subset(sc, Sales > 0 & Customers > 0))

```
The data falls along a straight line, whose slope we can figure out
using linear-regression,

```{r lmsalescustomers}

lm.sc <- lm(log10(Sales) ~ log10(Customers),
            data=subset(sc, Customers > 0 & Sales > 0))
summary(lm.sc)

```

Though the fit is not great, sales are sublinear in the number of
customers! Sales per customer fall with the number of customers!

There are dates in the data, that we extract years, months, weekdays
etc from

##Distribution
The distribution is log-normal --- a very clear and excellent
log-normal (not shown). What about distribution of sales at individual
stores?

```{r saleDistbnGivenStore}

distbnGivenStore <- function(store, trdata=training) {
    ys <- trdata$Sales[trdata$Store == store]
    ys <- ys[ys > 0]
    histogram(log10(ys))
}

```



## Baseline models
A very simple model would be to use the mean of all sales as the
predicted outcome. This model will have the standard-deviation of
the data as error,

```{r meanpred}

training.pos <- subset(training, Open == 1 & Sales > 0)
meanPred <- mean(log(training.pos$Sales))
err.mp <- sd(log(training.pos$Sales))

```

A more detailed model will use the means of a given store.

```{r meanpredbystore}

meanPred.store <- ddply(training.pos, .(Store), summarise,
                        count = length(Sales),
                        mean = mean(log(Sales)),
                        var = var(log(Sales)),
                        sd = sd(log(Sales)))

mvar <- sum(meanPred.store$var*meanPred.stor$count)/sum(meanPred.store$count)

print(paste("Error when the mean for each store is used as a prediction",
           sqrt(mvar)))


```
