`m``{r setup, include=FALSE}

knitr::opts_chunk$set(cache=FALSE)
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(include=TRUE)
knitr::opts_chunk$set(results="asis")
knitr::opts_chunk$set(fig.width=12)
knitr::opts_chunk$set(fig.height=14)
prjpath <- "~/work/learn/competitions/kaggle/rossmann/"
datapath <- paste(prjpath, "data/", sep="")
analpath <- paste(prjpath, "analysis/", sep="")
rcodepath <- paste(analpath, "Rcode/", sep="")
setwd(analpath)

```


```{r loadLibsAndSource, include=FALSE}

reqdpkgs <- c("lattice", "latticeExtra", "reshape2",
              "plyr",  "lubridate", "xgboost")
lapply(reqdpkgs, library, character.only=TRUE)

```


#Data for XGBoost

XGboost will not tolerate factor or character data, or NAs. We will
transform these variables,

```{r dataforxgb}

preds <- c("Store", "DayOfWeek", "Open", "Promo", "StateHoliday",
           "SchoolHoliday", "Year", "Month", "Week", "Mday", "Wday",
           "StoreType", "Assortment", "IsCompetition", "CompDuration",
           "Promo2", "Promo2DurationSinceRestart")

training.xgb <- training[, preds] #remove Sales and Customers
training.xgb <- do.call(cbind, lapply(training.xgb,
                                      function(xs) {
                                          if (class(xs) %in% c("factor",
                                                               "character")) {
                                              xs <- as.character(xs)
                                              xs[is.na(xs)] <- "missing"
                                              xs <- as.integer(as.factor(xs))
                                              xs
                                          } else {
                                              xs[is.na(xs)] <- -99999999999999
                                              xs[is.infinite(xs)] <- 99999999999999
                                              xs
                                          }
                                      })
                        )


```


From the scrolls we have seen a best of 0.116 test error. We should be
able to improve on this.

##Deviation from mean as outcome
Given that the out come is log-normal, lets try to use deviation from
the mean as a outcome.

```{r devout}

training$devmean <- with(training,
                         log(1 + Sales) - mean(log(1 + Sales), na.rm=TRUE))

```


# Independent categories

We should also try independent categories for the categorical variables.

```{r indpcats}

training.xgb <- training[, -c(4,5)]
training.xgb$StateHoliday <- as.integer(training.xgb$StateHoliday)

